---
title: "Report Analysis Birthweight"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 3
    number_sections: yes
  # word_document:
  #   toc: yes
  #   toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r global_options, include=FALSE,message=FALSE}
used.data <- databern %>%
       filter(multiple==0) %>%
       filter(!(weight < 1500)) %>%
       filter(!(matage > 50))  %>%
       filter(!(matage <14)) %>%
   filter(!(gest <30)) %>%
  mutate(
   Weight_quan = as.character(cut(weight, breaks=quantile(weight,probs=c(0,0.1,0.9,1)),include.lowest = TRUE)),
   Weight_cat_quan  = Weight_quan,
   Weight_cat_quan  = replace( Weight_cat_quan, Weight_cat_quan =="[1.5e+03,2.53e+03]","<10th percentile (< 2530 g)" ),
   Weight_cat_quan  = replace( Weight_cat_quan, Weight_cat_quan =="(2.53e+03,3.78e+03]","10th–90th percentile (2530 - 3780 g)" ),
   Weight_cat_quan  = replace( Weight_cat_quan, Weight_cat_quan =="(3.78e+03,5.58e+03]",">90th percentile (>3780 g)" ),
   Weight_cat_quan = factor(Weight_cat_quan, levels = c(">90th percentile (>3780 g)", 
                                                        "10th–90th percentile (2530 - 3780 g)",
                                                        "<10th percentile (< 2530 g)")),
   year = as.factor(year),
   boy = as.factor(boy),
   stillborn = as.factor(stillborn),
   city = as.factor(city),
   insurance = as.factor(insurance),
   married = as.factor(married),
   matbody2 = as.factor(matbody2),
   occupation2 = as.factor(occupation2),
   matheight2  = as.factor(matheight2 ),
   malnutrition2 = as.factor(malnutrition2),
   Gest_group = factor(Gest_group, levels=c("normal","early")),
   parity = factor(parity, levels=c("1","2","3",">=4")),
   city = factor(city, levels=c("1","0")),
   married = factor(married, levels=c("1","0")),
   boy = factor(boy, levels=c("1","0")),
   birth_season = factor( birth_season, levels=c("Spring","Summer","Fall","Winter")),
   parity = factor(parity, levels = c("1","2","3",">=4")),
   matbody2 = factor(matbody2, levels=c("2","1","3")),
   matheight2 = factor(matheight2, levels=c("2","1","3")),
   malnutrition2 = factor(malnutrition2, levels=c("0","1")),
   occupation2 = factor(occupation2, levels=c("4","1","2","3","5","6","7")),
   Exposure_sum_dummy = as.factor(Exposure_sum_dummy)
  )
```

# Birth weights

## Time series

No significant changes in birth weight 1918 and 1919. But birth weights increase (breakpoint) in May 1919 (week 27).
<br />
<br />
Weekly
<br />
<br />
```{r ,echo=FALSE, message=FALSE,warning=FALSE,fig.height = 10, fig.width = 10}
function_plot_kw(Timespan="weekly")
```

Monthly
<br />
<br />
```{r ,echo=FALSE, message=FALSE,warning=FALSE,fig.height = 10, fig.width = 10}
function_plot_kw(Timespan="monthly")
```


## unadjusted

The quantile regression does not provide any further information. The results are similar when using the quantile and linear regressions. No significant differences between the quantiles are found for the years 1918 and 1919. 
<br />
<br />
Lineare regression 
<br />
<br />
```{r, echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5 ,fig.width = 5}
function_quantile_regression_1914_1922(varExp="unadjusted_year_linear")
```

Quantile regression  
<br />
<br />
```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 8, fig.width = 8}
function_quantile_regression_1914_1922(varExp="unadjusted_year_qr")
```

Lineare regression only Exposure
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_quantile_regression_1914_1922(varExp="linear_Exp")
```

Lineare regression only flu intensity
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_quantile_regression_1914_1922(varExp="linear_Int")
```

GAM regression 
<br />
<br />
```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 8}
function_quantile_regression_1914_1922(varExp="unadjusted_gam_model")
```

GAM regression and Exposure 
<br />
<br />
```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height =5, fig.width =5}
function_quantile_regression_1914_1922(varExp="adjusted_gam_model_Exp")
```


GAM regression and flu intensity
<br />
<br />
```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height =5, fig.width =8}
function_quantile_regression_1914_1922(varExp="adjusted_gam_model_Int")
```

## adjusted 
Lineare regression  
<br />
<br />

Exposure sum
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_quantile_regression_1914_1922(varExp="adjusted_year_linear_Exp")
```

Flu Intensity
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_quantile_regression_1914_1922(varExp="adjusted_year_linear_Int")
```

GAM regression  without gestational age
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_quantile_regression_1914_1922(varExp="adjusted_gam_model")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 8}
function_quantile_regression_1914_1922(varExp="adjusted_gam_model_plot")
```

GAM regression  with gestational age
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_quantile_regression_1914_1922(varExp="adjusted_gam_model_gest")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 8}
function_quantile_regression_1914_1922(varExp="adjusted_gam_model_plot_gest")
```


# Gestational age
## Table

```{r Tableflu, echo=FALSE,results = "asis",message=FALSE}
used.data %>%
  mutate(year_num = as.integer(as.character(year)),
         year=as.character(year)) %>%
  filter(year_num >1913) %>%
   filter(stillborn==0) %>%
   group_by(year, Gest_group) %>%
   count() %>%
   ungroup() %>%
   spread(., Gest_group, n) %>%
    mutate(total = normal + early,
           perc_early = round((early/total)*100,2))%>%
   kbl() %>%
  kable_material(c("striped", "hover"),full_width = F,position = "left")
```



## Logistic Regression
I did the analysis with and without stillborns -> similar results. The results are shown with data without Stillborn.
High multicollinearity between year and exposure sum and year and flu intensity. Therefor only one variable is used.

unadjusted
<br />
<br />

```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_regression_gestage(varExp="unadjusted")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 5}
function_regression_gestage(varExp="unadjusted_plot")
```

only exposure
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_regression_gestage(varExp="unadjusted_Exp" )
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 5}
function_regression_gestage(varExp="unadjusted_Exp_plot" )
```

only flu intensity
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_regression_gestage(varExp="unadjusted_Int")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 5}
function_regression_gestage(varExp="unadjusted_Int_plot")
```

adjusted
<br />
<br />
year
```{r, echo=FALSE,message=TRUE,warning=FALSE}
function_regression_gestage(varExp="adjusted_year")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 8, fig.width = 5}
function_regression_gestage(varExp="adjusted_year_plot")
```

Exposure sum
```{r, echo=FALSE,message=TRUE,warning=FALSE}
function_regression_gestage(varExp="adjusted_Exp")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 8, fig.width = 5}
function_regression_gestage(varExp="adjusted_Exp_plot")
```

Flu_intensity_all
```{r, echo=FALSE,message=TRUE,warning=FALSE}
function_regression_gestage(varExp="adjusted_Int")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 8, fig.width = 5}
function_regression_gestage(varExp="adjusted_Int_plot")
```


# Stillborn
```{r, echo=FALSE,results = "asis",message=FALSE}
used.data %>%
  mutate(year_num = as.integer(as.character(year))) %>%
  filter(year_num >1911) %>%
  group_by(year,stillborn) %>%
  summarise(N=n())%>%
  filter(!is.na(stillborn)) %>%
  spread(., stillborn, N) %>%
  mutate(alive_d = `0`,
         stillborn_d = `1`,
         percentage_stillborn_d = round((stillborn_d/(alive_d+stillborn_d))*100 , 2)) %>%
   select(year, alive_d,stillborn_d,percentage_stillborn_d) %>%
   left_join(data.bfs.stillborn) %>%
   rename(alive_bfs = alive,
          stillborn_bfs = stillborn,
          perc_stillborn_bfs2 = percentage_stillborn,
          alive = alive_d,
          stillborn = stillborn_d,
          perc_stillborn = percentage_stillborn_d) %>%
   mutate(perc_stillborn_bfs = round((stillborn_bfs/(alive_bfs+stillborn_bfs))*100 , 2)) %>%
   select(- perc_stillborn_bfs2) %>%
      mutate(Perc_diff = perc_stillborn_bfs - perc_stillborn) %>%
       kbl(caption = "Data Bern") %>%
    kable_paper("hover", full_width = F)
```


## Logistic Regression

High multicollinearity between year and exposure sum and year and flu intensity. Therefor only one variable is used.

unadjusted
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_regression_stillborn(varExp="unadjusted")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 5}
function_regression_stillborn(varExp="unadjusted_plot")
```

Only Exposure
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_regression_stillborn(varExp="unadjusted_Exp")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 5}
function_regression_stillborn(varExp="unadjusted_Exp_plot")
```

Only Flu Intensity
<br />
<br />
```{r, echo=FALSE, message=TRUE,warning=FALSE}
function_regression_stillborn(varExp="unadjusted_Int")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 5, fig.width = 5}
function_regression_stillborn(varExp="unadjusted_Int_plot")
```

adjusted year
<br />
<br />
```{r, echo=FALSE,message=TRUE,warning=FALSE}
function_regression_stillborn(varExp="adjusted_year")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 8, fig.width = 5}
function_regression_stillborn(varExp="adjusted_year_plot")
```

adjusted Exp
<br />
<br />
```{r, echo=FALSE,message=TRUE,warning=FALSE}
function_regression_stillborn(varExp="adjusted_Exp")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 8, fig.width = 5}
function_regression_stillborn(varExp="adjusted_Exp_plot")
```


adjusted Int
<br />
<br />
```{r, echo=FALSE,message=TRUE,warning=FALSE}
function_regression_stillborn(varExp="adjusted_Int")
```

```{r , echo=FALSE, message=FALSE,warning=FALSE,fig.height = 8, fig.width = 5}
function_regression_stillborn(varExp="adjusted_Int_plot")
```